---
title: "SFMT2018"
author: "Yoshihiro Ishida"
date: '2018-09-03'
output:
  html_document:
    df_print: paged
---

```{r message=F}
library(tidyverse)
library(lubridate)

source("scripts/time-funs.R")
load("interim/sfmt.RData")
```


## 110km time chart based on 2015 result

__precaution__
In 2015, Otomi-ko aid (15.1km after Kurohime) was operated instead of Sasagamine (11km after Kurohime). Therefore, any prediction at Otomi is 30-40 minutes behind what would be expected in the current race.

```{r}
sfmt2015 <- sfmt2015 %>% filter(!rank2015 %in% c("TO", "DNF", "DNS")) %>% mutate(class100k = cut(finish, breaks = seq(9.5, 22.5, by = 1) * 3600, label = paste0(10:22, "h")))

sfmt2015_summary <- sfmt2015 %>% group_by(class100k) %>% summarize_at(vars(kurohime:finish), median, na.rm = T)
sfmt2015_clock <- sfmt2015_summary %>% mutate_at(vars(kurohime:finish), '+', 3600 * 5.5) %>% 
  mutate_at(vars(kurohime:finish), pretty_time)
sfmt2015_clock
```

```{r eval = F}
write_csv(sfmt2015_clock, path = "output/sfmt-100k-pred.csv")
```

## 100 mile time chart generated from convoluted method

1. extract runners who did both 2015 110km and 2017 100mile
2. merge 2017 result (Start - Kurohime) and 2015 resulst (Kurohime - finish)

```{r} 
sfmt2015_diff <- sfmt2015 %>% select(name_kana, kurohime:finish) %>% 
  mutate_at(vars(otomi:finish), '-', .$kurohime) %>% select(-kurohime)

both <- inner_join(sfmt2017, sfmt2015_diff, by = 'name_kana') %>% 
  mutate_at(vars(otomi:finish), '+', .$kurohime) %>% 
  filter(!is.na(kurohime) & rank2017 != "TO") %>% 
  mutate(total_time = cut(finish, breaks = seq(19.5, 32.5, by = 1) * 3600, labels = c(paste0(20:32, 'h')))) %>% 
  filter(!is.na(total_time))

mile_start <- as.POSIXct("2018-09-15 19:30:00")

both_stat <- both %>% group_by(total_time) %>% summarize_at(vars(banff:finish), median, na.rm = T)
both_clock <- both_stat %>% 
  mutate_at(vars(banff:finish), as.difftime, units = 'secs') %>% 
  mutate_at(vars(banff:finish), '+', mile_start)

both_clock

both_clock2 <- both_clock %>% mutate_at(vars(banff:finish), as.character, format = '%m-%d %H:%M')
```

```{r eval = F}
write_csv(both_clock2, "output/sfmt-100m-pred.csv")
```

